### 2023-10-02 chapters/background.tex
Work on discussion

\paragraph{Self-supervised learning}
An alternative approach to representation learning is provided by self-supervised learning \parencite{mikolov_efficient_2013,devlin_bert_2018,chen_simple_2020,schneider_wav2vec_2019} which is a form of unsupervised learning where the training objective is derived from the data itself. 
While it is fair to say that not all self-supervised training objectives are as well-motivated as the ELBO is for VAEs, they have shown impressive results within the field of natural language processing \parencite{devlin_bert_2018,chen_simple_2020}, speech processing \parencite{schneider_wav2vec_2019}, and computer vision \parencite{chen_simple_2020}. 
For a deeper comparison of VAEs and self-supervised methods for speech processing, we refer the reader to our paper \textcite{borgholt_brief_2022} included in \cref{app:paper-brief}. 
\Cref{chp:paper-review} provides a comprehensive introduction to and review of self-supervised learning in speech recognition.

>>>>

\paragraph{Self-supervised learning}
An alternative approach to representation learning is provided by self-supervised learning \parencite{mikolov_efficient_2013,devlin_bert_2018,chen_simple_2020,schneider_wav2vec_2019} which is a form of unsupervised learning where the training objective is derived from the data itself. 
While it is fair to say that not all self-supervised training objectives are as well-motivated as the ELBO is for VAEs, they have shown impressive results within the fields of natural language processing \parencite{devlin_bert_2018,chen_simple_2020}, speech processing \parencite{schneider_wav2vec_2019}, and computer vision \parencite{chen_simple_2020}. 
For a deeper comparison of VAEs and self-supervised methods for speech processing, we refer the reader to our paper \textcite{borgholt_brief_2022} included in \cref{app:paper-brief}. 
\Cref{chp:paper-review} provides a comprehensive introduction to and review of self-supervised learning in speech recognition.

### 2023-10-02 chapters/discussion.tex
Work on discussion

\section{\Cref{chp:paper-retrospective} revisited: Uncertainty}
The work in \textcite{wenstrup_retrospective_2023} focuses on predictive performance and an analysis of feature importance but does not explicitly consider uncertainty estimation.

\begin{figure}
    \begin{subfigure}[c]{0.48\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{python_plotting/calibration_curve_ensemble_and_all_models_uncalibrated.pdf}
        % \caption{}
        % \label{fig_discussion:calibration_curve_ensemble_and_all_models_uncalibrated}
    \end{subfigure}
    % \hfill
    \begin{subfigure}[c]{0.48\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{python_plotting/histogram_ensemble_and_single_model.pdf}
        % \caption{}
        % \label{fig_discussion:histogram_ensemble_and_single_model}
    \end{subfigure}
    \caption[Calibration curve for the uncalibrated stroke recognition model and empirical distribution of predicted probabilities.]{Calibration curve for the uncalibrated ensemble model with median F1-score also reported in \cref{tab_retrospective:table3-occlusion-analysis, fig_retrospective:figure1-roc-curve} (left) and the histogram of predicted probabilities (right).}
\end{figure}

>>>>

\section{\Cref{chp:paper-retrospective} revisited: Uncertainty}
The work in \textcite{wenstrup_retrospective_2023} focuses on the predictive performance of the ensemble model and an analysis of feature importance, but does not explicitly consider uncertainty estimation. 
As we discussed in \cref{subsec:model-calibration}, such a model can be calibrated to predict probabilities that are aligned with the empirical probability of the model being correct on some validation set.

In \cref{fig_discussion:retrospective-paper-calibration-curve-of-uncalibrated-model}, we plot the calibration curve for the uncalibrated stroke recognition ensemble and the individual models along with a histogram of its predicted probabilities. 
The calibration curve is computed by sorting the predicted probabilities on test set into a number of bins spanning the range from zero to one. For each bin, we compute the fraction of examples for which the model predicted correctly and then plot these pairs of mean predicted probabilities and correctness fractions. A perfectly calibrated model would have the same fraction of correct predictions in any given bin, as that bin's mean value. \Cref{fig_discussion:retrospective-paper-calibration-curve-of-uncalibrated-model} clearly shows the miscalibration issue that we previously discussed. 
It is interesting to note that the ensemble model is better calibrated than the individual models that form it. Since an ensemble output probability is computed as the harmonic mean of the five individual model probabilities, it is easy to see that the ensembles output probability can never exceed the maximum probability predicted among the individual models. This property tends to make ensemble probabilities less extreme and, since the individual models are overconfident, this results in better calibration (see also the histogram in \cref{fig_discussion:retrospective-paper-calibration-curve-of-uncalibrated-model}).

To calibrate the ensemble model, we can use methods such as Platt-scaling \parencite{platt_probabilistic_1999} or isotonic regression \parencite{zadrozny_transforming_2002}. 
In both cases, we fit a simple regression model (logistic or isotonic) to the predicted probabilities and the target labels on some validation set. 
In \cref{fig_discussion:retrospective-paper-calibration-curve-sigmoid-isotonic} we have done so for the ensemble model using our validation set for fitting and visualizing the results on the test set. On the left, we plot the resulting calibration curves and on the right we show the logistic and isotonic fits. We see that both methods result in quite good calibrations\footnote{Brier scores: Uncalibrated = $0.003500$, logistic = $0.001807$, isotonic = $0.001774$. Brier skill score compared to uncalibrated (relative improvement): Logistic = $0.4830$, isotonic = $0.4924$} and that the predicted probabilities are shifted towards smaller values.

However, while correct calibration provides some interpretability to the predictions, the resulting probabilities are still functions of the model's performance. For example, with the logistic calibration, we can note that the predicted probabilities have been shifted so far towards smaller values that the best the model can do, is a predicted probability of about 50\%. This means strokes can be predicted at best with 50\% probability of being correct. 
If this model were to be deployed in a randomized controlled trial, the calibrated probabilities might be useful to incorporate in the design of the user interface such that a user can discern between confident and uncertain model predictions. Nonetheless, it is easy to imagine how this generally low confidence might lead the user to ignore the model.

% Brier scores:  {'uncalibrated': 0.0034955788687128925, 'logistic': 0.0018072483306197486, 'isotonic': 0.0017744177396100578, 'average': 0.0021955477869598783}
% BSS uncalibrated reference:  {'logistic': 0.48299025755205005, 'isotonic': 0.4923822902432705}
% BSS mean reference:  {'logistic': 0.17685766561145932, 'isotonic': 0.1918109229282361}

\begin{figure}
    \begin{subfigure}[c]{0.48\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{python_plotting/calibration_curve_ensemble_and_all_models_uncalibrated.pdf}
        % \caption{}
        % \label{fig_discussion:calibration_curve_ensemble_and_all_models_uncalibrated}
    \end{subfigure}
    % \hfill
    \begin{subfigure}[c]{0.48\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{python_plotting/histogram_ensemble_and_single_model.pdf}
        % \caption{}
        % \label{fig_discussion:histogram_ensemble_and_single_model}
    \end{subfigure}
    \caption[Calibration curve for the uncalibrated stroke recognition model and empirical distribution of predicted probabilities.]{%
        Calibration curve for the uncalibrated stroke recognition ensemble model (left) and the histogram of predicted probabilities (right). 
        We use the ensemble that achieved the median F1-score which is the one also reported in \cref{tab_retrospective:table3-occlusion-analysis, fig_retrospective:figure1-roc-curve}.
    }
    \label{fig_discussion:retrospective-paper-calibration-curve-of-uncalibrated-model}
\end{figure}

\begin{figure}
    \begin{subfigure}[c]{0.48\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{python_plotting/calibration_curves_ensemble.pdf}
        % \caption{}
        % \label{fig_discussion:calibration_curve_ensemble_and_all_models_uncalibrated}
    \end{subfigure}
    % \hfill
    \begin{subfigure}[c]{0.48\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{python_plotting/calibration_fits_ensemble.pdf}
        % \caption{}
        % \label{fig_discussion:histogram_ensemble_and_single_model}
    \end{subfigure}
    \caption[Calibration fits and curves for the stroke recognition model using Platt-scaling and isotonic regression for calibration.]{%
        Calibration curves using sigmoid and isotonic calibration fits for the stroke recognition ensemble model (left) and the calibration fits (right). 
        We use the ensemble that achieved the median F1-score which is the one also reported in \cref{tab_retrospective:table3-occlusion-analysis, fig_retrospective:figure1-roc-curve}.
    }
    \label{fig_discussion:retrospective-paper-calibration-curve-sigmoid-isotonic}
\end{figure}

### 2023-10-02 chapters/introduction.tex
Work on discussion

The evolution of technology to support and automate decision-making can be traced back to the development of algorithms and mathematical techniques by ancient civilizations. 
The Egyptian Rhind Mathematical Papyrus, dated around 1550 BCE, describes various mathematical techniques for doing calculations involving multiplication, division, and fractions, and how to use them for practical purposes like measuring land, constructing buildings, and sharing resources \parencite{georges_universal_2001}. 
Around the same time, the Babylonians had developed a sophisticated system of mathematics that included algorithms for solving quadratic equations, calculating areas of shapes, and numerical methods for approximating square roots \parencite{fowler_square_1998}.

@@

\subsection{Model calibration} \label{subsec:model-calibration}
% 
The problem of miscalibration in machine learning models is well-known and has been studied for decades \parencite{lewis_sequential_1995, platt_probabilistic_1999, garczarek_classification_2002, zadrozny_transforming_2002, bennett_using_2003}. One of the best known methods for calibrating a binary classifier is Platt scaling which fits a logistic regression model to the model outputs, assuming that the miscalibration can be corrected by a logarithmic function \parencite{platt_probabilistic_1999}. Another method is isotonic regression which instead fits a non-parametric, monotonic function \parencite{zadrozny_transforming_2002}. 
More recently, \textcite{guo_calibration_2017} proposed a single-parameter variant of Platt-scaling that fits only a temperature parameter on the logits of a neural network classifier. 
These methods are generally simple and effective, but not without limitations. 
To perform the calibration, most of the methods require a held-out validation set on which the model was not trained, Platt scaling and isotonic regression are not naively applicable to multi-class classification problems, and how to calibrate models for structured prediction tasks, such as speech recognition and machine translation, remains an open problem \parencite{astudillo_uncertainty_2010, astudillo_integration_2013, jayashankar_detecting_2020}.

@@

\todo[inline]{Formulate research hypotheses}

This introduction provided a high-level overview of the motivation behind the research project by focusing on the use-cases of stroke recognition and automated medical coding as well as speech processing and the importance of uncertainty quantification. 
The remainder of \cref{part:background} consists of two chapters. 
\Cref{chp:main-contributions} outlines the research questions of the thesis via the published papers and presents their main contributions. 
\Cref{chp:technical-background} provides relevant technical background to the papers of the thesis, including uncertainty, out-of-distribution detection, and variational autoencoders.

\Cref{part:unsupervised-uncertainty-estimation,part:self-supervised-speech-representation-learning,part:medical-applications} contain a number of chapters that each correspond to a primary paper. 
\Cref{part:unsupervised-uncertainty-estimation} contains papers on variational autoencoders and their applications to out-of-distribution detection and speech modelling \parencite{havtorn_hierarchical_2021,havtorn_benchmarking_2022,bergamin_modelagnostic_2022}. 
\Cref{part:self-supervised-speech-representation-learning} contains a review paper on self-supervised speech representation learning \parencite{mohamed_selfsupervised_2022}. \Cref{part:medical-applications} contains papers on applications of machine learning within the medical domain, including recognition of stroke cases in calls to medical helplines \parencite{wenstrup_retrospective_2023} and medical coding of clinical notes \parencite{edin_automated_2023}.

>>>>

The evolution of technology to support and automate decision-making can be traced back to the development of algorithms and mathematical techniques by ancient civilizations. 
The Egyptian Rhind Mathematical Papyrus, dated to around 1550 BCE, describes various mathematical techniques for doing calculations involving multiplication, division, and fractions, and how to use them for practical purposes like measuring land, constructing buildings, and sharing resources \parencite{georges_universal_2001}. 
Around the same time, the Babylonians had developed a sophisticated system of mathematics that included algorithms for solving quadratic equations, calculating areas of shapes, and numerical methods for approximating square roots \parencite{fowler_square_1998}.

@@

\subsection{Model calibration} \label{subsec:model-calibration}
% 
The problem of miscalibration in machine learning models is well-known and has been studied for decades \parencite{lewis_sequential_1995, platt_probabilistic_1999, garczarek_classification_2002, zadrozny_transforming_2002, bennett_using_2003, niculescu-mizil_predicting_2005}. One of the best known methods for calibrating a binary classifier is Platt scaling which fits a logistic regression model to the model outputs, assuming that the miscalibration can be corrected by a logarithmic function \parencite{platt_probabilistic_1999}. Another method is isotonic regression which instead fits a non-parametric, monotonic function \parencite{zadrozny_transforming_2002}. 
More recently, \textcite{guo_calibration_2017} proposed a single-parameter variant of Platt-scaling that fits only a temperature parameter on the logits of a neural network classifier. 
These methods are generally simple and effective, but not without limitations. 
To perform the calibration, most of the methods require a held-out validation set on which the model was not trained, Platt scaling and isotonic regression are not naively applicable to multi-class classification problems, and how to calibrate models for structured prediction tasks, such as speech recognition and machine translation, remains an open problem \parencite{astudillo_uncertainty_2010, astudillo_integration_2013, jayashankar_detecting_2020}.

@@

This introduction provided a high-level overview of the motivation behind the research project by focusing on the use-cases of stroke recognition and automated medical coding as well as speech processing and the importance of uncertainty quantification. 
The remainder of \cref{part:background} consists of two chapters. 
\Cref{chp:main-contributions} presents the research questions of the thesis and the main contributions made in the included papers. 
\Cref{chp:technical-background} provides relevant technical background to the included papers, including uncertainty, out-of-distribution detection, and variational autoencoders.

\Cref{part:unsupervised-uncertainty-estimation,part:self-supervised-speech-representation-learning,part:medical-applications} contain a number of chapters that each correspond to a primary paper. 
\Cref{part:unsupervised-uncertainty-estimation} is made up of three papers on variational autoencoders and their applications to out-of-distribution detection and speech modelling \parencite{havtorn_hierarchical_2021,havtorn_benchmarking_2022,bergamin_modelagnostic_2022}. 
\Cref{part:self-supervised-speech-representation-learning} is a single, comprehensive review paper on self-supervised speech representation learning \parencite{mohamed_selfsupervised_2022}. 
\Cref{part:medical-applications} consists of two papers on applications of machine learning within the medical domain, specifically recognition of stroke cases in calls to medical helplines \parencite{wenstrup_retrospective_2023} and medical coding of clinical notes \parencite{edin_automated_2023}.

### 2023-10-02 chapters/main-contributions.tex
Work on discussion

The following chapters of the thesis are self-contained studies and therefore detail their own research questions and contributions, but do so without consideration to the other chapters. 
For that reason, we here lay out the research questions and main contributions of each chapter in relation to the overall research project of the thesis. 
The chapters are structured into three parts: {\color{dtured}\cref{part:unsupervised-uncertainty-estimation}} is concerned with unsupervised uncertainty estimation, {\color{dtured}\cref{part:self-supervised-speech-representation-learning}} is concerned with self-supervised speech representation learning, and {\color{dtured}\cref{part:medical-applications}} contains studies on machine learning methods applied for tasks in a medical setting. 
Finally, {\color{dtured}\cref{part:discussion-and-conclusion}} contains a discussion and the conclusion of the thesis.

>>>>

The following chapters of the thesis are self-contained studies and therefore detail their own research questions and contributions, but do so without consideration to the other chapters. 
For that reason, we here lay out the research questions and main contributions of each chapter in relation to the overall research project of the thesis. 
The chapters are structured into three parts: {\color{black}\cref{part:unsupervised-uncertainty-estimation}} is concerned with unsupervised uncertainty estimation, {\color{black}\cref{part:self-supervised-speech-representation-learning}} is concerned with self-supervised speech representation learning, and {\color{black}\cref{part:medical-applications}} contains studies on machine learning methods applied for tasks in a medical setting. 
Finally, {\color{black}\cref{part:discussion-and-conclusion}} contains a discussion and the conclusion of the thesis.
